---
title: "Midterm Part 3: Open Ended Analysis"
author: "Brandon Le"
output:
  prettydoc::html_pretty:
    theme: quartz
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(here)

min_wage <- read_csv(here("Midterm", "Minimum Wage Data.csv"))
```

# How do the minimum wages differ by region?

We'll group our states into regions classified by the United States Census: `Northeast`, `Midwest`, `South`, and `West`. Territories were not included in the US Census regions, but are included by their shortest distance to a particular region.


Let's take a look at the yearly minimum wage values(rescaled to 2020 dollars), averaged across each region.

```{r}
# defining our regions
# manually chose where Guam, Puerto Rico, and U.S. Virgin Islands will go

west = c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington", "Guam")

south = c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "District of Columbia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas", "Puerto Rico", "U.S. Virgin Islands")

midwest = c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")

northeast = c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")

# dummy code to show why I know these are the only territories:
# forgotten_regions <- min_wage %>%
# filter(!State %in% west & !State %in% south & !State %in% midwest & !State %in% northeast)
# forgotten_regions returns Guam, Puerto Rico, and U.S. Virgin Islands in the States Column 
# unless we include them in our regions.
```

```{r}
# computing the average minimum wage for each region for each year
min_region <- min_wage %>%
  select(Year, State, Effective.Minimum.Wage.2020.Dollars) %>%
  group_by(Year) %>%
  summarize(West = median(Effective.Minimum.Wage.2020.Dollars[State %in% west]),
            South = median(Effective.Minimum.Wage.2020.Dollars[State %in% south]),
            Midwest = median(Effective.Minimum.Wage.2020.Dollars[State %in% midwest]),
            Northeast = median(Effective.Minimum.Wage.2020.Dollars[State %in% northeast])              ) 
```

```{r}
#pivotlonging to make it easier to plot
min_plot <- pivot_longer(
              data = min_region, 
              cols = c(West:Northeast),
              names_to = "Region",
              values_to = "Average_Min_Wage_2020_Dollars"
            )


# plotting the average minimum wages for each year
ggplot(min_plot, 
       mapping = aes(
        x = Year, 
        y = Average_Min_Wage_2020_Dollars,
        color = Region
       ) 
) +
  geom_line() +
  geom_point(mapping = aes(shape = Region)) +
  labs(
    title = "Average(Median) Minimum Wage over Time",
    x = "Year",
    y = "Average Minimum Wage in 2020 Dollars") +
  theme(legend.position = "bottom")
```

Looking at this graph, we see that starting in 1968, The `Northeast` region was leading in average minimum wage when rescaled to 2020 dollars. However for the next few decades all the way until the 2000s the average minimum wage was effectively the same across all regions! In fact there seems to be a common factor creating a significant drop in effective minimum wage for all regions in the 1980s. After the year 2000, the `Northeast`'s average minimum wage rose significantly compared to the other regions until early 2010, and again after 2015. The `West` had the 2nd biggest spike after 2015, followed by the `Midwest`, while the `South`'s effective minimum wage seems to be continuously dropping. 

Now let's take a look at the summary statistics of the average minimum wage for all regions, across all years.
```{r}
summary(min_region)
```

The `Northeast` region has higher summary values across all fields, which matches our conclusion that the `Northeast`'s average minimum wage at least tied for the lead throughout the years.